---
title: "Practical2"
author: "Marc Bourleau"
date: "2023-10-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

(A)  Read in the data. Extract and represent the yearly max values from 1940 to 2009. What do you
observe ?

```{r}
library(ismev)
library(VGAM)
library(tidyverse)
library(plotly)
library(plotrix) 
library(scales)
```


```{r}
#QUESTION A

#Read the Data
venice <- venice90

venice_max <- venice %>%
  group_by(year) %>%
  summarise(max_sea_level = max(sealevel))

head(venice_max)

plot_ly(venice_max, x = ~year, y = ~max_sea_level, type = 'scatter', mode = 'markers', name = 'Max Value') %>% layout(title = "Maximum Value per year", xaxis = list(title="Year"), yaxis = list(title="Maximum Value"))  %>% add_segments(x = 1940, xend = 2009, y = 140, yend = 140, line = list(color = 'red', width = 2))

```

We will be using the Peak over threshold method for this part. From the plot, we can see that there are a total of 11 outliers in the data over the threshold of 140 (red line), with 1966 being the highest value overall.


```{r}
#QUESTION B

# making predictions with linear models (remainder)
mod1 <- lm(max_sea_level ~ year, data = venice_max)
summary(mod1)

# making predictions for 5 years
mod1_predict <- predict.lm(mod1,newdata=data.frame("year"=c(2010:2022)),se=T, interval = "confidence", level = 0.95)
mod1_predict

```

Globally, from the confidence intervals, the values range from 124 to 146, which doesn't provide a very good fit for the model. This may be due to the lack of historical data to accompany the model.



```{r}
#QUESTION C

#Add the model in a data frame
venice_max_predict <- data.frame(
  PredictedValues = mod1_predict
) %>% 
  mutate(year = c(2010:2022))
venice_max_predict 

#Plot the confidence intervals
plotCI(x = venice_max_predict$year,
       y = venice_max_predict$PredictedValues.fit.fit,
       li = venice_max_predict$PredictedValues.fit.lwr,
       ui = venice_max_predict$PredictedValues.fit.upr)

#Create a new dataframe for the extreme values of 2010 - 2022
max_real <- data.frame(year = c(2012, 2012, 2013, 2018, 2019, 2019, 2019, 2022), max_sea_level = c(143, 149, 143, 156, 187, 144, 154, 204))


ggplot() +
  geom_ribbon(data = venice_max_predict, aes(x = year, ymin = PredictedValues.fit.lwr, ymax = PredictedValues.fit.upr), fill = "blue", alpha = 0.3) +
  geom_point(data = max_real, aes(x = year, y = max_sea_level, color = "Observed", colour = "red"), alpha = 0.5) +
  geom_point(data = venice_max_predict, aes(x = year, y = PredictedValues.fit.fit, color = "Predicted"), shape = 1) +
  labs(title = "Real extreme values from 2010 to 2022 versus predicted values' confidence intervals", x = "Year", y = "Sea Level") +
  scale_x_continuous(breaks = unique(c(venice_max_predict$year, max_real$year))) +
  scale_color_manual(name = "Data Type", values = c("Observed" = "red", "Predicted" = "black")) 




```
ASK WETHER WE SHOULD INCLUDE ALL DATA POINTS FROM THE WIKIPEDIA TABLE OR NOT

We can see that the predicted values are not representative of the observed values. The only value that is well-predicted is for the 13th of November 2019, as the Confidence Interval includes the actual value. 


(d) Fit a GEV a with constant parameters to the historical yearly max values. Fit a GEV with time varying location parameter. Compare the two embedded models using likelihood ratio test (LRT). Show diagnostic plots.
```{r}
library(extRemes)
list_max_sea_levels <- unlist(venice_max$max_sea_level)
gev_fix <- fevd(list_max_sea_levels, type = "GEV", time.units= "years")
plot(gev_fix) ; gev_fix$results$par
??fevd
model <- gev.fit(venice_max$max_sea_level ~ venice_max$year, type = "location", data = venice_max)
??ismev::gev.fit
gev_time_varying <- fevd(venice_max, type = "GEV", time.units = "years", formula = max_sea_level ~ year)
model <- gev.fit(venice_max$max_sea_level ~ venice_max$year, type = "location", data = venice_max)

time_varying_gev_model <- gev.fit(venice_max$max_sea_level ~ venice_max$year, data = venice_max)

mod3 <- fevd(rain.ann.max, type="GEV", time.units="years")
plot(mod3)
mod3$results$par #gives parameters of the GEV
```

