---
title: "Practical2.part3"
output: html_document
date: "2023-11-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, message = FALSE, warning=FALSE}
source(here::here("script/setup.R"))
```
# Part 3 
## (a) Read in the data for the daily night maximum temperatures in Lausanne. Subset the summer months (June to September).
```{r question 3a}
night_max <- read_csv(here::here("data/nightmax.csv"))  
night_min <- read_csv(here::here("data/nightmin.csv"))
# Remove rows with missing values
night_max <- na.omit(night_max)
night_min <- na.omit(night_min)
# Subset the data for summer months (June to September)
summer_night_max <- night_max[format(night_max$date,"%m") %in% c("06", "07", "08", "09"), ] 
```

## (b) Assess whether extremes of the subsetted series in (a) occur in cluster.
```{r question 3b}
# Visualise the data: plot and histogram of the maximum temperatures for the summer months 
plot(summer_night_max$date, summer_night_max$night.max, type = "l", xlab = "Date", ylab = "Night Max Temperature", main = "Summer Night Max Temperatures")
hist(summer_night_max$night.max, breaks = 30, xlab = "Night Max", ylab = "Frequency", main = "Histogram of summer night max temperatures")
# Have more information about the distribution 
min(summer_night_max$night.max) ; mean(summer_night_max$night.max); max(summer_night_max$night.max)
quantile(summer_night_max$night.max, 0.95) # doesn't seem to be very right-skewed 
# Choose the threshold using the mrlplot()
mrlplot(summer_night_max$night.max, main="Mean residual")
threshrange.plot(summer_night_max$night.max, r= c(20,30), nint = 20) # we are going to use 30 
th_1 <- 30 # with this threshold we would have 5% of observations: 400 so seems good. 
# Visualise the threshold 
plot(night_max$night.max, type = 'l')
abline(h=th,col=2) # looks good 
# Assess the threshold
pot_mle <- fevd(night_max$night.max, method = "MLE", type="GP", threshold=th_1) 
plot(pot_mle) 
rl_mle <- return.level(pot_mle, conf = 0.05, return.period= c(2,5,10,20,50,100), do.ci=T) # diagnostic plot looks good and coinfidence interval become wider the longer the return level, which make sense 
# Return level plots with MLE 
par(mfcol=c(1,1))
plot(pot_mle, type="rl",
     main="Return Level Plot for Oberwang w/ MLE",
    ylim=c(0,200), pch=16)
loc <- as.numeric(return.level(pot_mle, conf = 0.05, return.period=50))
segments(50, 0, 50, loc, col= 'midnightblue',lty=6)
segments(0.01,loc,50, loc, col='midnightblue', lty=6)
# Assess wether extremes occur in a cluster using extremalindex()
extremalindex(night_max$night.max, th_1) # 0.31 so 1/0.31 = 3.23 WE NEED TO DECLUSTER 


```
We have an approximate cluster size of 3.23, indicating that extremes values are clustered together. Therefore, we need to decluster. 

## (c) Decluster the data from (a) using a suitable threshold. Plot the resulting declustered data. (Hint: you may want to use the extRemes package.)
```{r question 3c}
# We need to create a vector of size night_max that stores the year. 
years <- numeric(nrow(night_max))

k <- 1
for (i in 1:nrow(night_max)) {
  if (is.na(night_max$night.max[i])) {
    next
  } else {
    years[k] <- year(as.Date(night_max$date[i]))
    k <- k + 1
  }
}
years <- years-1999 # we don't want the year but the index of the years
# Use decluster function of extRemes 
decl1 <- decluster(night_max$night.max, threshold=th, groups=years, na.action=na.omit) # vector and groups need to have the same size 
decl1 # we have 71 clusters 
plot(decl1) # shows in grey the points that are not retained in the selection 
```

## (d) Fit a GPD to the data, both raw and declustered. Assess the quality of the fit. 
```{r}
# Use fevd of the normal and declustered 
gpd_raw <- fevd(night_max$night.max, threshold = th_1, type = "GP")
gpd_declustered <- fevd(decl1, threshold = th_1, type = "GP")
plot(gpd_raw)
plot(gpd_declustered)
```

(e) Repeat the above analysis for the negatives of the daily nightly minimum temperatures for the
winter months (November-February).
```{r question 3e}
# Subset the winter months (November to February)
winter_night_min <- night_min[format(night_min$date, "%m") %in% c("11", "12", "01", "02"), ]
# We first do the plot and histogram of the negatives of the daily nightly minimum temperatures for the winter months (November-February).
plot(winter_night_min$date, winter_night_min$night.min, type = "l", xlab = "Date", ylab = "Night Min Temperature", main = "Winter Night Min Temperatures")
hist(winter_night_min$night.min, breaks = 30, xlab = "Date", ylab = "Night Min Temperature", main = "Winter Night Min Temperatures frequency")
min(winter_night_min$night.min); mean(winter_night_min$night.min); max(winter_night_min$night.min)
quantile(winter_night_min$night.min, 0,95)

mrlplot(winter_night_min$night.min, main="Mean residual")
threshrange.plot(winter_night_min$night.min, r= c(-5,15), nint =20)

mean(winter_night_min$night.min)

max(summer_night_max$night.max)
quantile(summer_night_max$night.max, 0,95)
mean(summer_night_max$night.max)

```

