---
title: "TP3_RA"
output: html_document
date: "2023-11-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
source(here::here("script/setup.R"))
```

(a) Read in the data and select only the winter temperature data. You can use the code lines of the file GetData.R to get the data and select only the winter values.
```{r}
# You may use the follwoing packages for EVT analysis
library(extRemes)
library(ismev)

##### Download daily NAO measurements
library(data.table)
NAO.daily <- 
  fread('ftp://ftp.cdc.noaa.gov/Public/gbates/teleconn/nao.reanalysis.t10trunc.1948-present.txt')
NAO.daily <- as.matrix(NAO.daily)
colnames(NAO.daily) <- c("year","month","day","NAO")

##### Download temperature data: be sure to work in the correct directory
months <- c(12,1,2) #keep only winter observations

library(readr)
temp_max_Zermatt <- read_delim("daily_maximum_Zermatt/order_107669_data.txt", 
                                         ";", escape_double = FALSE, col_types = 
                                           cols(time = col_number()), 
                                         trim_ws = TRUE, skip = 1)

colnames(temp_max_Zermatt) <- c("station","time","temp")
temp_max_Zermatt           <- temp_max_Zermatt[-1,]
temp_max_Zermatt[,2]       <- as.Date(apply(temp_max_Zermatt[,2],1,as.character),
                                      "%Y%m%d")

temp_max_Montana <- read_delim("daily_maximum_Montana/order_107668_data.txt", ";", escape_double = FALSE, col_types = cols(time = col_number()), trim_ws = TRUE, skip = 1)

colnames(temp_max_Montana) <- c("station","time","temp")
temp_max_Montana           <- temp_max_Montana[-1,]
temp_max_Montana[,2]       <- as.Date(apply(temp_max_Montana[,2],1,as.character),
                                      "%Y%m%d")

###match the dates of the two time series
temp_max_Montana           <- temp_max_Montana[match(as.matrix(temp_max_Zermatt[,2]),
                                                     as.matrix(temp_max_Montana[,2])),]
temp_max_Montana           <- as.matrix(temp_max_Montana)
colnames(temp_max_Montana) <- c("station","time","temp")
temp_max_Zermatt           <- as.matrix(temp_max_Zermatt)
colnames(temp_max_Zermatt) <- c("station","time","temp")

###keep only winter dates
library(lubridate)
temp_max_Montana <- temp_max_Montana[which(month(as.POSIXlt(temp_max_Montana[,"time"], 
                                                            format="%Y-%m-%d")) %in% months),]
temp_max_Zermatt <- temp_max_Zermatt[which(month(as.POSIXlt(temp_max_Zermatt[,"time"], 
                                                            format="%Y-%m-%d")) %in% months),]


Date       <- function( length = 0 ){
  newDate = numeric( length )
  class(newDate) = "Date"
  return(newDate)
}

season_day                   <- yday(as.Date(temp_max_Montana[,2]))
season_day[season_day < 61]  <- season_day[season_day < 61] + 31
season_day[season_day > 334] <- season_day[season_day > 334]- 334

NAO.date <- Date(nrow(NAO.daily))
for(i in 1:nrow(NAO.daily)){
  NAO.date[i] <- as.Date(paste(as.character(NAO.daily[i,1]),"-",as.character(NAO.daily[i,2]),
                               "-",as.character(NAO.daily[i,3]),sep=""),format="%Y-%m-%d")
}
nao  <- NAO.daily[intersect(as.Date(temp_max_Montana[,2]),as.Date(NAO.date)),4]


#Montana
x_Montana          <- data.frame("time"=temp_max_Montana[,2],
                                 "nao"=nao,
                                 "d"=season_day,
                                 "temp"=temp_max_Montana[,3])

as.numeric.factor  <- function(x) {as.numeric(levels(x))[x]}
x_Montana[,"temp"] <- as.numeric(x_Montana[,"temp"])
x_Montana[,"time"] <- as.numeric(year(as.POSIXlt(x_Montana[,"time"], format="%Y-%m-%d")))
x_Montana[,"time"] <- (x_Montana[,"time"]-min(x_Montana[,"time"]))/(max(x_Montana[,"time"])-
                                                                      min(x_Montana[,"time"]))

#Zermatt
x_Zermatt          <- data.frame("time"=temp_max_Zermatt[,2],
                                 "nao"=nao,
                                 "d"=season_day,
                                 "temp"=temp_max_Zermatt[,3])

as.numeric.factor  <- function(x) {as.numeric(levels(x))[x]}
x_Zermatt[,"temp"] <- as.numeric(x_Zermatt[,"temp"])
x_Zermatt[,"time"] <- as.numeric(year(as.POSIXlt(x_Zermatt[,"time"], format="%Y-%m-%d")))
x_Zermatt[,"time"] <- (x_Zermatt[,"time"]-min(x_Zermatt[,"time"]))/(max(x_Zermatt[,"time"])-
                                                                      min(x_Zermatt[,"time"]))

Z <- data.frame("Montana"=x_Montana[,"temp"] , "Zermatt"=x_Zermatt[,"temp"], "NAO"=x_Zermatt[,"nao"])
```


(b) Represent the data and especially scatterplots of Zermatt values against Montana’s values. Same with Montana values against NAO values. Comment.

(c) Test the correlation between these two pairs of series.

(d) Now concentrate on the extreme level and tail dependence. Represent χ and χ ̄ for two pairs of series (Montana vs Zermatt and Montana vs NAO). You may want to use the function chiplot of the package evd. What do you observe?